### 2015-06-23 Correspondence/revisions.tex
added response letter; added some responses to the manuscript. changes the code which generates the table in the paper.

\documentclass[a4paper,10pt]{letter}
\usepackage[utf8]{inputenc}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{Laura White\\Editorial Board Member\\Scientific Reports }
\address{Escola de Matem\'atica Aplicada\\Funda\c{c}\~ao Getulio Vargas 
(FGV)\\Rio de Janeiro -- RJ\\Brazil.}

\opening{Dear editor,}

In response to the comments of the reviewers of our manuscript, we have made 
the following modifications to the manuscript:

\begin{description}
 \item[Reviewer 1:] responses are listed below.

 \begin{enumerate}
  \item A simulation experiment was done, as requested and included as a new 
section in the supplementary material. The simulations shows that the inference 
methodology can recover the exact parameters user to simulate  the data.
\item The text of the methods section was improved to answer the four issues 
raised in this comment.
 \end{enumerate}

 \item [Reviewer 2:]
\end{description}

\signature{Flávio Codeço Coelho\\Professor}

\closing{Sincerely}

%enclosure listing
%\encl{}

\end{letter}
\end{document}