### 2015-07-06 Correspondence/revisions.tex
Done responding to comments 2 and 3 of reviewer 1 and described the changes in the letter. Closes #22 Closes #23

\documentclass[a4paper,10pt]{letter}
\usepackage[utf8]{inputenc}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{Laura White\\Editorial Board Member\\Scientific Reports }
\address{Escola de Matem\'atica Aplicada\\Funda\c{c}\~ao Getulio Vargas 
(FGV)\\Rio de Janeiro -- RJ\\Brazil.}

\opening{Dear editor,}

In response to the comments of the reviewers of our manuscript, we have made 
the following modifications to the manuscript:

\begin{description}
 \item[Reviewer 1:] Responses to the main comments are listed below. The minor 
comments where also accepted, namely the addition of readme do the code 
repository as well as the citation of two  additional papers.

 \begin{enumerate}
  \item A simulation experiment was done, as requested and included as a new 
section in the supplementary material. The simulations shows that the inference 
methodology can recover the exact parameters user to simulate  the data.
\item The text of the methods section was improved to answer the four issues 
raised in this comment.
\item The introduction and discussion sections were expanded to included the 
additional information required by this comment.
 \end{enumerate}

 \item [Reviewer 2:] \textbf{Minor comments:} Fixed typo in the abstract; 
 \begin{enumerate}
  \item 
 \end{enumerate}

\end{description}

\signature{Flávio Codeço Coelho\\Professor}

\closing{Sincerely}

%enclosure listing
%\encl{}

\end{letter}
\end{document}