### 2015-07-06 Correspondence/revisions.tex
finished responding to reviewer 2's comments. Closes #27, closes #28, closes #37

\documentclass[a4paper,10pt]{letter}
\usepackage[utf8]{inputenc}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{Laura White\\Editorial Board Member\\Scientific Reports }
\address{Escola de Matem\'atica Aplicada\\Funda\c{c}\~ao Getulio Vargas 
(FGV)\\Rio de Janeiro -- RJ\\Brazil.}

\opening{Dear editor,}

In response to the comments of the reviewers of our manuscript, we have made 
the following modifications to the manuscript:

\begin{description}
 \item[Reviewer 1:] Responses to the main comments are listed below. The minor 
comments where also accepted, namely the addition of readme do the code 
repository as well as the citation of two  additional papers.

 \begin{enumerate}
  \item A simulation experiment was done, as requested and included as a new 
section in the supplementary material. The simulations shows that the inference 
methodology can recover the exact parameters user to simulate  the data.
\item The text of the methods section was improved to answer the four issues 
raised in this comment.
\item The introduction and discussion sections were expanded to included the 
additional information required by this comment.
 \end{enumerate}

 \item [Reviewer 2:] \textbf{Minor comments:} Fixed typo in the abstract; Added 
text to the discussion stating that we expect the effects of underreporting to 
cancel out in the calculation of the Attack ratio. 
 \begin{enumerate}
  \item Reference 5, despite having similar concepts as those discussed in our 
paper, has quite different goals. we have however, cited another publication 
(ref 26), which fro the same group and based on the same field study, and use 
their findings about seroprevalence in the 'results and discussion' section.
\item We have not found articles combining the use of aggregated data to study 
variable force of infection, however we have reviewed other works modeling 
variable FoI (refs 7, 8, 9, 10, 12).
 \end{enumerate}

\end{description}

\signature{Flávio Codeço Coelho\\Professor}

\closing{Sincerely}

%enclosure listing
%\encl{}

\end{letter}
\end{document}