### 2015-08-31 maps_figure.tex
first stab at making the new maps figure

% \documentclass{standalone}
% \usepackage{capt-of}
% \usepackage{subfigure}
% \usepackage{graphicx}
% \begin{document}
% %  \begin{center}
% %  \includegraphics[width=8cm,height=4.82cm]{./plots/heatmap2010.jpg}
%   \subfigure[2010]{\includegraphics[width=8cm,height=4.82cm]{./plots/heatmap2010.jpg}}
%   \subfigure[2011]{\includegraphics[width=8cm]{./plots/heatmap2011.jpg}}
%   \subfigure[2012]{\includegraphics[width=8cm]{./plots/heatmap2012.jpg}}
%   \subfigure[2013]{\includegraphics[width=8cm]{./plots/heatmap2013.jpg}}
% %  \end{center}
% %    \captionof{figure}{}
% \end{document}

\documentclass[preview]{standalone}
\usepackage{mwe}
\usepackage[]{subfig}
\usepackage[margin=0.1cm]{geometry}
\begin{document}
% \def\FigSize{.23}
\begin{figure}[H]
\centering
\subfloat[\textbf{2010}]
  {
 \includegraphics[width=0.485\textwidth,height=4.0cm]{./plots/heatmap2010.jpg}
  }
\subfloat[\textbf{2011}]
  {
 \includegraphics[width=0.485\textwidth,height=4.0cm]{./plots/heatmap2011.jpg}
  }\\
\subfloat[\textbf{2012}]
  {
 \includegraphics[width=0.485\textwidth,height=4.0cm]{./plots/heatmap2012.jpg}
  }
  \subfloat[\textbf{2013}]
  {
 \includegraphics[width=0.485\textwidth,height=4.0cm]{./plots/heatmap2013.jpg}
  }
\end{figure}
\end{document}